# Preface {.unnumbered}

- Holds code resources for class exercise

## Necessary packages

```{r, echo = F}
# load tibble, dplyr, naniar
library(tibble)
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(naniar))
suppressPackageStartupMessages(library(readr))
suppressPackageStartupMessages(library(kableExtra))
suppressPackageStartupMessages(library(ggplot2))
```

If you do not have these installed ("Could not find package xyz") you need to 
install the package. 

```{r, eval = F}
# load necessary packages
# data preparation
library(readr)
library(tibble)
library(dplyr)

# visualization
library(ggplot2)

# missings
library(naniar)

# tables
library(kableExtra)
```

## Import dataset

- By storing it into an object with the `zz_`-prefix, you can easier modify it 
  afterwards without re-importing

```{r}
# import the dataset using the {readr}-package
zz_dataset <- read_csv("data/2023-02-20_QoC_sample_dataset.csv", 
                       show_col_types = FALSE)
```


## Check the import

By looking at the top five rows (the "head" of the dataset): 

```{r}
head(zz_dataset, n = 5)
```

- Seems the input worked as expected :relieved:
- the columns 'sex' and 'unit' should be a factor, not a character

```{r}
# change 'sex' into factor
dataset <- 
  zz_dataset %>% 
  mutate(sex = as.factor(sex), 
         unit = as.factor(unit))
```


... then again check the top five rows:


```{r}
head(dataset, n = 5)
```